---
title: Analysis
author: ak
format: 
  html:
    self-contained: true
    
    code-tools: true 
    source: true
    fig-width: 10
    fig-height: 6
---

```{r}
#Load libraries
library(tidyverse)
library(lme4)
library(ggfortify)
library(lmerTest)
library(performance)
library(readr)
#install.packages("broom.mixed")
library(broom.mixed)
library(grid)
library(ggeffects)
#install.packages("sjPlot")
library(sjPlot)
library(sjlabelled)
```

```{r}
droughtnet <- read_csv("droughtnet_data_cleaned_final.csv", 
                 col_types = cols(calluna_shoot_type = col_factor(levels = c("Long", "Short"))),
                 guess_max = Inf)

View(droughtnet)
# Now, to reorder the ageClass factor levels to start with Pioneer
droughtnet$ageClass <- factor(droughtnet$ageClass, levels = c("Pioneer", "Mature"))
levels(droughtnet$ageClass)

```


```{r}
droughtnet2 <- droughtnet %>%
  mutate(plant_nr = case_when(
    # For Calluna vulgaris under Amb (0) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Calluna vulgaris" & 
    DroughtTrt == "Amb (0)" ~ paste0(siteID, "_", ageClass, "_CV_Amb(0)_", DroughNet_plotID, "_", plant_nr),

    # For Calluna vulgaris under Ext (90) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Calluna vulgaris" & 
    DroughtTrt == "Ext (90)" ~ paste0(siteID, "_", ageClass, "_CV_Ext(90)_", DroughNet_plotID, "_", plant_nr),

    # For Vaccinium myrtillus under Amb (0) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Vaccinium myrtillus" & 
    DroughtTrt == "Amb (0)" ~ paste0(siteID, "_", ageClass, "_VM_Amb(0)_", DroughNet_plotID, "_", plant_nr),

    # For Vaccinium myrtillus under Ext (90) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Vaccinium myrtillus" & 
    DroughtTrt == "Ext (90)" ~ paste0(siteID, "_", ageClass, "_VM_Ext(90)_", DroughNet_plotID, "_", plant_nr),

    # For Vaccinium vitis-idaea under Amb (0) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Vaccinium vitis-idaea" & 
    DroughtTrt == "Amb (0)" ~ paste0(siteID, "_", ageClass, "_VV_Amb(0)_", DroughNet_plotID, "_", plant_nr),

    # For Vaccinium vitis-idaea under Ext (90) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Vaccinium vitis-idaea" & 
    DroughtTrt == "Ext (90)" ~ paste0(siteID, "_", ageClass, "_VV_Ext(90)_", DroughNet_plotID, "_", plant_nr),

    # For Empetrum nigrum under Amb (0) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Empetrum nigrum" & 
    DroughtTrt == "Amb (0)" ~ paste0(siteID, "_", ageClass, "_EN_Amb(0)_", DroughNet_plotID, "_", plant_nr),

    # For Empetrum nigrum under Ext (90) treatment
    siteID %in% c("Lygra", "Tjøtta") & 
    ageClass %in% c("Pioneer", "Mature") & 
    species == "Empetrum nigrum" & 
    DroughtTrt == "Ext (90)" ~ paste0(siteID, "_", ageClass, "_EN_Ext(90)_", DroughNet_plotID, "_", plant_nr),

    TRUE ~ as.character(plant_nr)
  ))
view(droughtnet2)
```


```{r}
# Specify the columns to check for NA values
# List of columns to check for NA values
columns_to_check <- c("DroughtTrt", "DroughNet_plotID", "ageClass", "siteID", 
                      "species", "plant_height", "SLA", "mean_thickness", "dry_mass_g_original", "LDMC", "leaf_area", "wet_mass_g", "dry_mass_g")


# Check for rows with any NA values in the specified columns
na_rows <- apply(droughtnet2[columns_to_check], 1, function(row) any(is.na(row)))

# Create a data frame showing how many NA values there are in each row
na_row_summary <- droughtnet2[na_rows, columns_to_check]
na_row_summary$row_na_count <- rowSums(is.na(na_row_summary))

# Summarize the number of rows by the count of NAs they contain
na_row_summary <- na_row_summary %>%
  group_by(row_na_count) %>%
  summarise(n = n(), .groups = 'drop')

# Create a summary showing the co-occurrence of NA values across the specified columns
na_co_occurrence <- droughtnet2 %>%
  summarise(across(all_of(columns_to_check), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "na_count") %>%
  arrange(desc(na_count))

# Print the summaries
print(paste("Total rows with any NA values across checked columns:", sum(na_rows)))
print(na_row_summary)
print(na_co_occurrence)
```

```{r}
# Create the new data frame with selected columns
droughtnet_data2 <- select(droughtnet2, 
                           "envelope_ID", "siteID", "species", "ageClass", "DroughtTrt", 
                           "DroughNet_plotID", "leaf_age", "calluna_shoot_type", 
                           "plant_height", "mean_thickness", "SLA", "LDMC", "leaf_area", "plant_nr")

view(droughtnet_data2)

# Specify the columns to check for NA values
columns_to_check2 <- c("plant_height", "mean_thickness", "SLA", "LDMC", "leaf_area")

# Remove rows with any NA values in the specified columns
droughtnet_data2_clean <- droughtnet_data2[!rowSums(is.na(droughtnet_data2[columns_to_check2])), ]
view(droughtnet_data2_clean)
# Calculate the total number of NAs removed (for the specified columns only)
# This calculation considers rows removed rather than individual NAs
total_NAs_removed <- nrow(droughtnet_data2) - nrow(droughtnet_data2_clean)

# Calculate the total number of entries remaining
total_entries_remaining <- nrow(droughtnet_data2_clean)

# Print out the total NAs removed and total entries remaining
cat("Total NAs removed:", total_NAs_removed, "\n")
cat("Total entries remaining:", total_entries_remaining, "\n")
```

**Analysis for young leaves**

```{r}
# Specify the species of interest, including Calluna vulgaris
species_of_interest <- c("Empetrum nigrum", "Vaccinium myrtillus", "Vaccinium vitis-idaea", "Calluna vulgaris")

# Filter the data to include only the specified species where leaf_age is 'young',
# and for Calluna vulgaris, additionally check that calluna_shoot_type is 'short'

filtered_data <- droughtnet_data2_clean %>%
  filter(species %in% species_of_interest, leaf_age == "young") %>%
  filter(
    # Include Calluna vulgaris from Tjøtta regardless of calluna_shoot_type
    (species == "Calluna vulgaris" & siteID == "Tjøtta") |
    # Include Calluna vulgaris from Lygra only if calluna_shoot_type is Short
    (species == "Calluna vulgaris" & siteID == "Lygra" & calluna_shoot_type == "Short") |
    # Include other species of interest from both sites
    (species != "Calluna vulgaris")
  )
view(filtered_data)

```

**visualization of raw data**

**plant height**

```{r}
filtered_data  <- filtered_data  %>%
  filter(siteID %in% c("Tjøtta", "Lygra"),
         DroughtTrt %in% c("Amb (0)", "Ext (90)"),
         ageClass %in% c("Pioneer", "Mature"))

ggplot(filtered_data, aes(x = species, y = plant_height, fill = DroughtTrt)) +
  geom_boxplot() +
  facet_grid(ageClass ~ siteID, scales = "free") +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Plant Height vs Treatment by Species for Young Leaves",
       x = "Species",
       y = "Plant Height") +
  scale_fill_manual(values = c("Amb (0)" = "blue", "Ext (90)" = "grey"))  
```

**SLA**

```{r}
ggplot(filtered_data, aes(x = species, y = log(SLA), fill = DroughtTrt)) +
  geom_boxplot() +
  facet_grid(ageClass ~ siteID, scales = "free") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        plot.caption = element_text(hjust = 0.5)) + 
  labs(title = "SLA vs Treatment by Species",
       x = "Species",
       y = "Log (SLA)",
       fill = "Drought Treatment",
       caption = "Figure 2: SLA Variation in Dwarf Shrubs Under Drought Conditions Across Different Ages and Sites") +
  scale_fill_manual(values = c("Amb (0)" = "orange", "Ext (90)" = "darkgrey")) 
```

**LDMC**

```{r}
ggplot(filtered_data, aes(x = species, y = LDMC, fill = DroughtTrt)) +
  geom_boxplot() +
  facet_grid(ageClass ~ siteID, scales = "free") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        plot.caption = element_text(hjust = 0.5)) + 
  labs(title = "LDMC vs Treatment by Species",
       x = "Species",
       y = "LDMC",
       fill = "Drought Treatment",
       caption = "Figure 3: LDMC Variation in dwarf Shrubs Under drought conditions across different successional phases and sites") +
  scale_fill_manual(values = c("Amb (0)" = "darkblue", "Ext (90)" = "lightblue")) 


#trying using jitter for the graphs
ggplot(filtered_data, aes(x = species, y = LDMC, color = DroughtTrt)) +
  geom_jitter(position = position_dodge(width = 0.5), size = 1.5, alpha = 0.7) + # Use position_dodge to separate points
  facet_grid(ageClass ~ siteID, scales = "free") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        plot.caption = element_text(hjust = 0.5)) + 
  labs(title = "LDMC vs Treatment by Species",
       x = "Species",
       y = "LDMC",
       color = "Drought Treatment", # Updated from fill to color
       caption = "Figure 3: LDMC Variation in Dwarf Shrubs Under Drought Conditions Across Different Successional Phases and Sites") +
  scale_color_manual(values = c("Amb (0)" = "darkblue", "Ext (90)" = "red")) 

```


Analysis for combined data set
Using for way interactions to check if site is an influential factor in functional trait response to drought for each trait
```{r}
#LDMC
LDMC_d_s_a_r <- lmer(LDMC ~ DroughtTrt * species * ageClass * siteID +
                                    (1 | DroughNet_plotID), 
                                  data = filtered_data)

summary(LDMC_d_s_a_r)
anova(LDMC_d_s_a_r)


performance::check_model(LDMC_d_s_a_r, detrend = FALSE)


```

```{r}
#THICKNESS
thickness_d_s_a_r <- lmer(mean_thickness ~ DroughtTrt * species * ageClass * siteID +
                                    (1 | DroughNet_plotID), 
                                  data = filtered_data)

summary(thickness_d_s_a_r)
anova(thickness_d_s_a_r)


performance::check_model(thickness_d_s_a_r, detrend = FALSE)

```

```{r}
#HEIGHT
height_d_s_a_r <- lmer(log(plant_height) ~ DroughtTrt * species * ageClass * siteID +
                                    (1 | DroughNet_plotID), 
                                  data = filtered_data)

summary(height_d_s_a_r)
anova(height_d_s_a_r)


performance::check_model(height_d_s_a_r, detrend = FALSE)
```

```{r}
#SLA
sla_d_s_a_r <- lmer(log(SLA) ~ DroughtTrt * species * ageClass * siteID +
                                    (1 | DroughNet_plotID), 
                                  data = filtered_data)

summary(sla_d_s_a_r)
anova(sla_d_s_a_r)

#validate model
performance::check_model(sla_d_s_a_r, detrend = FALSE)
```

```{r}
#leaf area
area_d_s_a_r <- lmer(log(leaf_area) ~ DroughtTrt * species * ageClass * siteID +
                                    (1 | DroughNet_plotID), 
                                  data = filtered_data)

summary(area_d_s_a_r)
anova(area_d_s_a_r)

#validate model
performance::check_model(area_d_s_a_r, detrend = FALSE)
```
While "site" as a main factor for does not significantly affect the functional trait response to drought in some traits, its interactions with other factors are significant. This indicates that the effect of drought on the functional trait is strongly influenced by the site, and there are also significant three-way and four-way interactions involving "site." These findings suggest that the role of "site" is complex and becomes important when considered in combination with other factors like species, drought treatment, and age class.
The presence of significant interactions involving site underscores the importance of considering geographic location or environmental conditions in functional trait responses to drought.

We therefore proceed to partition the data by site


```{r}

# Function to format p-values
format_pvalue <- function(p) {
  if (p < 0.001) {
    "<0.001"
  } else {
    sprintf("%.3f", p)
  }
}

# Run ANOVA on each model, tidy the results, and format the p-values
anova_sla <- anova(sla_d_s_a_r) %>% tidy() %>% mutate(p.value = map_chr(p.value, format_pvalue)) %>% mutate(Trait = "SLA")
anova_height <- anova(height_d_s_a_r) %>% tidy() %>% mutate(p.value = map_chr(p.value, format_pvalue)) %>% mutate(Trait = "Plant height")
anova_thickness <- anova(thickness_d_s_a_r) %>% tidy() %>% mutate(p.value = map_chr(p.value, format_pvalue)) %>% mutate(Trait = "THICKNESS")
anova_LDMC <- anova(LDMC_d_s_a_r) %>% tidy() %>% mutate(p.value = map_chr(p.value, format_pvalue)) %>% mutate(Trait = "LDMC")
anova_area <- anova(area_d_s_a_r) %>% tidy() %>% mutate(p.value = map_chr(p.value, format_pvalue)) %>% mutate(Trait = "leaf_area")

# Combine the results into a single table
anova_combined <- bind_rows(anova_sla, anova_height, anova_thickness, anova_LDMC, anova_area)

# Use pivot_wider() to reshape the data from long to wide format
anova_wide <- anova_combined %>%
  select(term, Trait, p.value) %>%
  pivot_wider(names_from = Trait, values_from = p.value)

# Print the wide-format table
print(anova_wide)

write.csv(anova_wide, file = "anova_summary.csv", row.names = FALSE)

```


1 **Analysis for lygra site**

**filter data only for lygra**

```{r}
# Filter the data for siteID Lygra
lygra_data <- filtered_data %>%
  filter(siteID == "Lygra")
view(lygra_data)
lygra_data$species <- as.factor(lygra_data$species)
lygra_data$ageClass <- as.factor(lygra_data$ageClass)
lygra_data$DroughtTrt <- as.factor(lygra_data$DroughtTrt)

```

```{r}
ggplot(lygra_data, aes(x = species, y = leaf_area, fill = DroughtTrt)) +
  geom_boxplot() +
  facet_grid(~ ageClass, scales = "free") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        plot.caption = element_text(hjust = 0.5)) + 
  labs(title = "LDMC vs Treatment by Species",
       x = "Species",
       y = "leaf area",
       fill = "Drought Treatment",
       caption = "Figure 3: leaf area Variation in dwarf Shrubs Under drought conditions across different successional phases and sites") +
  scale_fill_manual(values = c("Amb (0)" = "blue", "Ext (90)" = "lightblue")) 

```


**`Trait 1- LDMC- models`**

```{r}
LY_LDMC_s_d_a <- lmer(LDMC ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = lygra_data)

summary(LY_LDMC_s_d_a)


#validate model
performance::check_model(LY_LDMC_s_d_a, detrend = FALSE)
anova(LY_LDMC_s_d_a)
emmeans(LY_LDMC_s_d_a, ~ DroughtTrt * species * ageClass )
#Tukey post-hoc test
#1. species effefct
emm_interaction1 <- emmeans(LY_LDMC_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_interaction1)

#2. drought effect
# For assessing the effect of drought across all species and age classes
emm_drought1 <- emmeans(LY_LDMC_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_drought1)

# 3. age effect
# For assessing the effect of age class within each species and drought treatment
emm_age1 <- emmeans(LY_LDMC_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_age1)

```

```{r}
#over lay the three way interaction on the plot for data 
if (!requireNamespace("emmeans", quietly = TRUE)) {
  install.packages("emmeans")
}
library(emmeans)

if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)

if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
library(ggplot2)

# Calculate estimated marginal means for the interaction
emm_ldmc <- emmeans(LY_LDMC_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_ldmc_df <- summary(emm_ldmc) %>%
  as.data.frame() %>%
  mutate(species = recode(species,
                          "Calluna vulgaris" = "CV",
                          "Vaccinium myrtillus" = "VM",
                          "Empetrum nigrum" = "EN",
                          "Vaccinium vitis-idaea" = "VV"))

# Define the dodge width for the boxplots and points
dodge_width <- 0.75

# Create a data frame for annotations
annotations <- data.frame(
  species = rep(c("CV", "EN", "VM", "VV"), times = 4),
  DroughtTrt = rep(c("Amb (0)", "Ext (90)"), each = 4),
  ageClass = rep(c("Pioneer", "Mature"), each = 8),
  stringsAsFactors = FALSE
)

# # Then, use dplyr to mutate the label based on conditions
annotations <- annotations %>%
  mutate(label = case_when(
    ageClass == "Pioneer" & species == "CV" ~ "e",
    ageClass == "Pioneer" & species == "EN" & DroughtTrt == "Amb (0)" ~ "f",
    ageClass == "Pioneer" & species == "EN" & DroughtTrt == "Ext (90)" ~ "g",
    ageClass == "Pioneer" & species == "VM" & DroughtTrt == "Amb (0)" ~ "h",
    ageClass == "Pioneer" & species == "VM" & DroughtTrt == "Ext (90)" ~ "i",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Amb (0)" ~ "j",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Ext (90)" ~ "k",
    ageClass == "Mature" & species == "CV" ~ "a",
    ageClass == "Mature" & species == "EN" ~ "b",
    ageClass == "Mature" & species == "VM" ~ "c",
    ageClass == "Mature" & species == "VV" ~ "d",
    TRUE ~ NA_character_  # default case to handle any unspecified combinations
  ))


# Calculate the y position for annotations
y_positions <- lygra_data %>%
  group_by(species, DroughtTrt, ageClass) %>%
  summarize(y_position = max(LDMC) * 1.05, .groups = "drop")  
# Merge annotation labels with their y positions
annotations <- merge(annotations, y_positions, by = c("species", "DroughtTrt", "ageClass"))

# Now let's create the plot
p <- ggplot(lygra_data, aes(x = species, y = LDMC, fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  geom_point(data = emm_ldmc_df, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, species, ageClass)),
             position = position_dodge(width = dodge_width), size = 1.5) +
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, size = 14, face = "bold"),
        axis.text.y = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14, face = "bold"),
        strip.text = element_text(size = 12, face = "bold")) +
  labs(title = "LDMC variability",
       x = "Species",
       y = "LDMC",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(name = "Drought Treatment", 
                    values = c("Amb (0)" = "blue", "Ext (90)" = "skyblue"),
                    labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  scale_color_manual(name = "Drought Treatment",
                     values = c("Amb (0)" = "salmon", "Ext (90)" = "salmon"),
                     labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  geom_text(data = annotations, aes(x = species, y = y_position, label = label),
            position = position_dodge(width = dodge_width), size = 3.5, fontface = "bold", vjust = -0.5)  
# Print the plot
print(p)
# Save the plot
ggsave("LDMC_plot_adjusted.png", plot = p, width = 7, height = 4, dpi = 300)

```


**Trait 2- SLA**

```{r}

LY_sla_s_d_a <- lmer(log(SLA) ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = lygra_data)
summary(LY_sla_s_d_a)
#validate model
performance::check_model(LY_sla_s_d_a, detrend = FALSE)
anova(LY_sla_s_d_a)
emmeans(LY_sla_s_d_a, ~ DroughtTrt * species * ageClass )
# tukey post-hoc
#1. species effefct
emm_interaction2 <- emmeans(LY_sla_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_interaction2)

#2. drought effect
# For assessing the effect of drought across all species and age classes
emm_drought2 <- emmeans(LY_sla_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_drought2)

# 3. age effect
# For assessing the effect of age class within each species and drought treatment
emm_age2 <- emmeans(LY_sla_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_age2)

```

```{r}
# Check and load required packages
if (!requireNamespace("emmeans", quietly = TRUE)) {
  install.packages("emmeans")
}
library(emmeans)

if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)

# Calculate estimated marginal means for the interaction
emm_sla <- emmeans(LY_sla_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_sla_df <- summary(emm_sla) %>%
  as.data.frame() %>%
  mutate(species = recode(species,
                          "Calluna vulgaris" = "CV",
                          "Vaccinium myrtillus" = "VM",
                          "Empetrum nigrum" = "EN",
                          "Vaccinium vitis-idaea" = "VV"))

# dodge width for the boxplots 
dodge_width <- 0.75

# Create a data frame for annotations
annotations <- data.frame(
  species = rep(c("CV", "EN", "VM", "VV"), times = 4),
  DroughtTrt = rep(c("Amb (0)", "Ext (90)"), each = 4),
  ageClass = rep(c("Pioneer", "Mature"), each = 8),
  stringsAsFactors = FALSE
)

#  mutate the label based on conditions
annotations <- annotations %>%
  mutate(label = case_when(
    ageClass == "Pioneer" & species == "CV" ~ "h",
    ageClass == "Pioneer" & species == "EN" ~ "i",
    ageClass == "Pioneer" & species == "VM" ~ "j",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Amb (0)" ~ "k",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Ext (90)" ~ "l",
    ageClass == "Mature" & species == "CV" ~ "a",
    ageClass == "Mature" & species == "EN" & DroughtTrt == "Amb (0)" ~ "b",
    ageClass == "Mature" & species == "EN" & DroughtTrt == "Ext (90)" ~ "c",
    ageClass == "Mature" & species == "VM" & DroughtTrt == "Amb (0)" ~ "d",
    ageClass == "Mature" & species == "VM" & DroughtTrt == "Ext (90)" ~ "e",
    ageClass == "Mature" & species == "VV" & DroughtTrt == "Amb (0)" ~ "f",
    ageClass == "Mature" & species == "VV" & DroughtTrt == "Ext (90)" ~ "g", 
    TRUE ~ NA_character_  
  ))

# Calculate the y position for annotations
y_positions <- lygra_data %>%
  group_by(species, DroughtTrt, ageClass) %>%
  summarize(y_position = max(log(SLA)) * 1.05, .groups = "drop") 
# Merge annotation labels with their y positions
annotations <- merge(annotations, y_positions, by = c("species", "DroughtTrt", "ageClass"))

p <- ggplot(lygra_data, aes(x = species, y = log(SLA), fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  geom_point(data = emm_sla_df, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, species, ageClass)),
             position = position_dodge(width = dodge_width), size = 1.5) +
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, size = 14, face = "bold"),
        axis.text.y = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14, face = "bold"),
        strip.text = element_text(size = 12, face = "bold")) +
  labs(title = "SLA variability",
       x = "Species",
       y = "log (SLA) cm²/g",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(name = "Drought Treatment", 
                    values = c("Amb (0)" = "blue", "Ext (90)" = "skyblue"),
                    labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  scale_color_manual(name = "Drought Treatment",
                     values = c("Amb (0)" = "salmon", "Ext (90)" = "salmon"),
                     labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  geom_text(data = annotations, aes(x = species, y = y_position, label = label),
            position = position_dodge(width = dodge_width), size = 3.5, fontface = "bold", vjust = 0)  

print(p)
# Save the plot
ggsave("sla_plot_adjusted.png", plot = p, width = 7, height = 4, dpi = 300)


```


**Trait 3: plant height**

```{r}
LY_h_s_d_a <- lmer(log(plant_height) ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = lygra_data)

summary(LY_h_s_d_a)
emmeans(LY_h_s_d_a, ~ DroughtTrt * species * ageClass )


#validate model
performance::check_model(LY_h_s_d_a, detrend = FALSE)
anova(LY_h_s_d_a)
#post hoc_test
#1. species effefct
emm_interaction <- emmeans(LY_h_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_interaction)

#2. drought effect
emm_drought3 <- emmeans(LY_h_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_drought3)

# 3. age effect
emm_age <- emmeans(LY_h_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_age)


```


```{r}
# Check and load required packages
if (!requireNamespace("emmeans", quietly = TRUE)) {
  install.packages("emmeans")
}
library(emmeans)

if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)

# Calculate estimated marginal means for the interaction
emm_h <- emmeans(LY_h_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_height_df <- summary(emm_h) %>%
  as.data.frame() %>%
  mutate(species = recode(species,
                          "Calluna vulgaris" = "CV",
                          "Vaccinium myrtillus" = "VM",
                          "Empetrum nigrum" = "EN",
                          "Vaccinium vitis-idaea" = "VV"))

dodge_width <- 0.75

# Create a data frame for annotations
annotations <- data.frame(
  species = rep(c("CV", "EN", "VM", "VV"), times = 4),
  DroughtTrt = rep(c("Amb (0)", "Ext (90)"), each = 4),
  ageClass = rep(c("Pioneer", "Mature"), each = 8),
  stringsAsFactors = FALSE
)

#mutate the label based on conditions
annotations <- annotations %>%
  mutate(label = case_when(
    ageClass == "Pioneer" & species == "CV" ~ "e",
    ageClass == "Pioneer" & species == "EN" ~ "f",
    ageClass == "Pioneer" & species == "VM" & DroughtTrt == "Amb (0)" ~ "g",
    ageClass == "Pioneer" & species == "VM" & DroughtTrt == "Ext (90)" ~ "h",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Amb (0)" ~ "i",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Ext (90)" ~ "j",
    ageClass == "Mature" & species == "CV" ~ "a",
    ageClass == "Mature" & species == "EN" ~ "b",
    ageClass == "Mature" & species == "VM" & DroughtTrt == "Amb (0)" ~ "c",
    ageClass == "Mature" & species == "VM" & DroughtTrt == "Ext (90)" ~ "d",
    ageClass == "Mature" & species == "VV" ~ "i",  
    TRUE ~ NA_character_  
  ))

# Calculate the y position for annotations
y_positions <- lygra_data %>%
  group_by(species, DroughtTrt, ageClass) %>%
  summarize(y_position = max(log(plant_height)) * 1.05, .groups = 'drop') 

# Merge annotation labels with their y positions
annotations <- merge(annotations, y_positions, by = c("species", "DroughtTrt", "ageClass"))

p <- ggplot(lygra_data, aes(x = species, y = log(plant_height), fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  geom_point(data = emm_height_df, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, species, ageClass)),
             position = position_dodge(width = dodge_width), size = 1.5) +
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, size = 14, face = "bold"),
        axis.text.y = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14, face = "bold"),
        strip.text = element_text(size = 12, face = "bold")) +
  labs(title = "Plant height variability",
       x = "Species",
       y = "log(Plant height) (cm)",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(name = "Drought Treatment", 
                    values = c("Amb (0)" = "blue", "Ext (90)" = "skyblue"),
                    labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  scale_color_manual(name = "Drought Treatment",
                     values = c("Amb (0)" = "salmon", "Ext (90)" = "salmon"),
                     labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  geom_text(data = annotations, aes(x = species, y = y_position, label = label),
            position = position_dodge(width = dodge_width), size = 3.5, fontface = "bold", vjust = 0)  
# Print the plot
print(p)

ggsave("height_plot_adjusted.png", plot = p, width = 7, height = 4, dpi = 300)

```


**Trait 4: mean -thickness

```{r}
LY_t_s_d_a <- lmer(mean_thickness ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = lygra_data)

summary(LY_t_s_d_a)

#validate model
performance::check_model(LY_t_s_d_a, detrend = FALSE)
anova(LY_t_s_d_a)
emmeans(LY_t_s_d_a, ~ DroughtTrt * species * ageClass )

# For example, for species within each DroughtTrt and ageClass
#1. species effefct
emm_interaction <- emmeans(LY_t_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_interaction)

#2. drought effect
# For assessing the effect of drought across all species and age classes
emm_drought <- emmeans(LY_t_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_drought)

# 3. age effect
# For assessing the effect of age class within each species and drought treatment
emm_age <- emmeans(LY_t_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_age)

```

overlay three way interaction

```{r}
# Check and load required packages
if (!requireNamespace("emmeans", quietly = TRUE)) {
  install.packages("emmeans")
}

if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)


# Calculate estimated marginal means for the interaction
emm_thickness <- emmeans(LY_t_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_thickness_df <- summary(emm_thickness) %>%
  as.data.frame() %>%
  mutate(species = recode(species,
                          "Calluna vulgaris" = "CV",
                          "Vaccinium myrtillus" = "VM",
                          "Empetrum nigrum" = "EN",
                          "Vaccinium vitis-idaea" = "VV"))

dodge_width <- 0.75

# Create a data frame for annotations
annotations <- data.frame(
  species = rep(c("CV", "EN", "VM", "VV"), times = 4),
  DroughtTrt = rep(c("Amb (0)", "Ext (90)"), each = 4),
  ageClass = rep(c("Pioneer", "Mature"), each = 8),
  stringsAsFactors = FALSE
)

annotations <- annotations %>%
  mutate(label = case_when(
    ageClass == "Pioneer" & species == "CV" ~ "e",
    ageClass == "Pioneer" & species == "EN" ~ "f",
    ageClass == "Pioneer" & species == "VM" ~ "g",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Amb (0)" ~ "h",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Ext (90)" ~ "e",
    ageClass == "Mature" & species == "CV" ~ "a",
    ageClass == "Mature" & species == "EN" ~ "b",
    ageClass == "Mature" & species == "VM" ~ "c",
    ageClass == "Mature" & species == "VV" ~ "d",  
    TRUE ~ NA_character_  
  ))

# Calculate the y position for annotations
y_positions <- lygra_data %>%
  group_by(species, DroughtTrt, ageClass) %>%
  summarize(y_position = max(mean_thickness) * 1.05, .groups = 'drop') # Adjust the multiplier as needed

# Merge annotation labels with their y positions
annotations <- merge(annotations, y_positions, by = c("species", "DroughtTrt", "ageClass"))

p <- ggplot(lygra_data, aes(x = species, y = mean_thickness, fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  geom_point(data = emm_thickness_df, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, species, ageClass)),
             position = position_dodge(width = dodge_width), size = 1.5) +
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, size = 14, face = "bold"),
        axis.text.y = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14, face = "bold"),
        strip.text = element_text(size = 12, face = "bold")) +
  labs(title = "Leaf thickness variability",
       x = "Species",
       y = "Leaf thickness (mm)",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(name = "Drought Treatment", 
                    values = c("Amb (0)" = "blue", "Ext (90)" = "skyblue"),
                    labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  scale_color_manual(name = "Drought Treatment",
                     values = c("Amb (0)" = "salmon", "Ext (90)" = "salmon"),
                     labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  geom_text(data = annotations, aes(x = species, y = y_position, label = label),
            position = position_dodge(width = dodge_width), size = 3.5, fontface = "bold", vjust = 0)  
print(p)
ggsave("thickness_plot_adjusted.png", plot = p, width = 7, height = 4, dpi = 300)

```

trait 5. leaf area
```{r}
#use log plus 1 transformation because leaf area has values which are almost zero
lygra_data$leaf_area_transformed <- log(lygra_data$leaf_area + 1)

LY_area_s_d_a <- lmer(leaf_area_transformed ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = lygra_data)

summary(LY_area_s_d_a)

#validate model
performance::check_model(LY_area_s_d_a, detrend = FALSE)
anova(LY_area_s_d_a)

#post hoc-yest
#1. species effefct
emm_interaction5 <- emmeans(LY_area_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_interaction5)

#2. drought effect
emm_drought5 <- emmeans(LY_area_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_drought5)

# 3. age effect
emm_age5 <- emmeans(LY_area_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_age5)
```


```{r}
# Check and load required packages
if (!requireNamespace("emmeans", quietly = TRUE)) {
  install.packages("emmeans")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)


# Calculate estimated marginal means for the interaction
emm_a <- emmeans(LY_area_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_area_df <- summary(emm_a) %>%
  as.data.frame() %>%
  mutate(species = recode(species,
                          "Calluna vulgaris" = "CV",
                          "Vaccinium myrtillus" = "VM",
                          "Empetrum nigrum" = "EN",
                          "Vaccinium vitis-idaea" = "VV"))

dodge_width <- 0.75

# Create a data frame for annotations
annotations <- data.frame(
  species = rep(c("CV", "EN", "VM", "VV"), times = 4),
  DroughtTrt = rep(c("Amb (0)", "Ext (90)"), each = 4),
  ageClass = rep(c("Pioneer", "Mature"), each = 8),
  stringsAsFactors = FALSE
)

annotations <- annotations %>%
  mutate(label = case_when(
    ageClass == "Pioneer" & species == "CV" ~ "f",
    ageClass == "Pioneer" & species == "EN" ~ "g",
    ageClass == "Pioneer" & species == "VM" & DroughtTrt == "Amb (0)" ~ "h",
    ageClass == "Pioneer" & species == "VM" & DroughtTrt == "Ext (90)" ~ "i",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Amb (0)" ~ "j",
    ageClass == "Pioneer" & species == "VV" & DroughtTrt == "Ext (90)" ~ "k",
    ageClass == "Mature" & species == "CV" ~ "a",
    ageClass == "Mature" & species == "EN" ~ "b",
    ageClass == "Mature" & species == "VM" & DroughtTrt == "Amb (0)" ~ "c",
    ageClass == "Mature" & species == "VM" & DroughtTrt == "Ext (90)" ~ "d",
    ageClass == "Mature" & species == "VV" ~ "e",  
    TRUE ~ NA_character_
  ))

# Calculate the y position for annotations
y_positions <- lygra_data %>%
  group_by(species, DroughtTrt, ageClass) %>%
  summarize(y_position = max(log(leaf_area)) * 1.05, .groups = 'drop') # Adjust the multiplier as needed

# Merge annotation labels with their y positions
annotations <- merge(annotations, y_positions, by = c("species", "DroughtTrt", "ageClass"))

p <- ggplot(lygra_data, aes(x = species, y = leaf_area_transformed, fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  geom_point(data = emm_area_df, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, species, ageClass)),
             position = position_dodge(width = dodge_width), size = 1.5) +
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, size = 14, face = "bold"),
        axis.text.y = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14, face = "bold"),
        strip.text = element_text(size = 12, face = "bold")) +
  labs(title = "Leaf area variability",
       x = "Species",
       y = "Leaf area (cm²)",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(name = "Drought Treatment", 
                    values = c("Amb (0)" = "blue", "Ext (90)" = "skyblue"),
                    labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  scale_color_manual(name = "Drought Treatment",
                     values = c("Amb (0)" = "salmon", "Ext (90)" = "salmon"),
                     labels = c("Amb (0)" = "Ambient", "Ext (90)" = "Extreme Drought")) +
  geom_text(data = annotations, aes(x = species, y = y_position, label = label),
            position = position_dodge(width = dodge_width), size = 3.5, fontface = "bold", vjust = -0.5)

print(p)
ggsave("area_plot_adjusted.png", plot = p, width = 7, height = 4, dpi = 300)

```

tabulating the anova summary
```{r}
extract_interaction_p <- function(anova_model) {
  term <- rownames(anova_model)[nrow(anova_model)]
  p_value <- anova_model[nrow(anova_model), "Pr(>F)"]
  c(term = term, p_value = p_value)
}

interaction_ldmc <- extract_interaction_p(anova_ldmc)
interaction_sla <- extract_interaction_p(anova_sla)
interaction_h <- extract_interaction_p(anova_h)
interaction_t <- extract_interaction_p(anova_t)

# Combine into a data frame
interactions_df <- data.frame(
  Trait = c("LDMC", "SLA", "Height", "Thickness"),
  Interaction_Term = c(interaction_ldmc["term"], interaction_sla["term"], interaction_h["term"], interaction_t["term"]),
  P_Value = c(interaction_ldmc["p_value"], interaction_sla["p_value"], interaction_h["p_value"], interaction_t["p_value"])
)
# save the table to a CSV file
write.csv(interactions_df, "ANOVA_Summary_Interactions.csv", row.names = FALSE)

```


**site: Tjøtta**

```{r}
# Filter the data for siteID tjøtta
tjøtta_data <- filtered_data %>%
  filter(siteID == "Tjøtta")
view(tjøtta_data)
```

**trait 1- LDMC**

```{r}
tj_LDMC_s_d_a <- lmer(LDMC ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = tjøtta_data)

summary(tj_LDMC_s_d_a)


#validate model
performance::check_model(tj_LDMC_s_d_a, detrend = FALSE)
anova(tj_LDMC_s_d_a)
#post hoc-test using emmeans
#1. species effefct
emm_tj_species <- emmeans(tj_LDMC_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_tj_species)

#2. drought effect
emm_tj_drought <- emmeans(tj_LDMC_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_tj_drought)

# 3. age effect
emm_tj_ageclass <- emmeans(tj_LDMC_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_tj_ageclass)
```

```{r}
if(!requireNamespace("emmeans", quietly = TRUE)) install.packages("emmeans")
if(!requireNamespace("pbkrtest", quietly = TRUE)) install.packages("pbkrtest")


# Calculate estimated marginal means for the interaction
emm_ldmc_tj <- emmeans(tj_LDMC_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_tj <- summary(emm_ldmc_tj) %>%
  as.data.frame()

# Define the position dodge width
dodge_width <- 0.75

# Plotting with boxplots for the raw data
p <- ggplot(tjøtta_data, aes(x = species, y = LDMC, fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  # Overlay the EMM predictions as points for clarity on the model's fit
  # Ensure that the position dodge here matches the dodge for the boxplot
  geom_point(data = emm_tj, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, ageClass)),
             position = position_dodge(width = dodge_width), size = 3) +
  # Facet by ageClass to compare across age classes
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "LDMC by Treatment, Species, and Age Class",
       x = "Species",
       y = "LDMC",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(values = c("Amb (0)" = "lightblue", "Ext (90)" = "salmon")) +
  scale_color_manual(values = c("Amb (0)" = "blue", "Ext (90)" = "red"))

# Display the plot
print(p)


```



**Trait2- plant height**

```{r}
tj_height_s_d_a <- lmer(log(plant_height) ~ DroughtTrt * species* ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = tjøtta_data)
summary(tj_height_s_d_a)


#validate model
performance::check_model(tj_height_s_d_a, detrend = FALSE)
anova(tj_height_s_d_a)

#1. species effefct
emm_tj_species2 <- emmeans(tj_height_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_tj_species2)

#2. drought effect
emm_tj_drought2 <- emmeans(tj_height_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_tj_drought2)

# 3. age effect
emm_tj_ageclass2 <- emmeans(tj_height_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_tj_ageclass2)
```


```{r}
if(!requireNamespace("emmeans", quietly = TRUE)) install.packages("emmeans")
if(!requireNamespace("pbkrtest", quietly = TRUE)) install.packages("pbkrtest")


# Calculate estimated marginal means for the interaction
emm_height_tj <- emmeans(tj_height_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_tj <- summary(emm_height_tj) %>%
  as.data.frame()

# Define the position dodge width
dodge_width <- 0.75

# Plotting with boxplots for the raw data
p <- ggplot(tjøtta_data, aes(x = species, y = log(plant_height), fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  # Overlay the EMM predictions as points for clarity on the model's fit
  # Ensure that the position dodge here matches the dodge for the boxplot
  geom_point(data = emm_tj, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, ageClass)),
             position = position_dodge(width = dodge_width), size = 3) +
  # Facet by ageClass to compare across age classes
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "LDMC by Treatment, Species, and Age Class",
       x = "Species",
       y = "Plant Height",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(values = c("Amb (0)" = "lightblue", "Ext (90)" = "salmon")) +
  scale_color_manual(values = c("Amb (0)" = "blue", "Ext (90)" = "red"))

# Display the plot
print(p)

```




**trait3 - SLA**

```{r}
tj_sla_s_d_a <- lmer(log(SLA) ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = tjøtta_data)
summary(tj_sla_s_d_a)


#validate model
performance::check_model(tj_sla_s_d_a, detrend = FALSE)
anova(tj_sla_s_d_a)


#1. species effefct
emm_tj_species3 <- emmeans(tj_sla_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_tj_species3)

#2. drought effect
emm_tj_drought3 <- emmeans(tj_sla_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_tj_drought3)

# 3. age effect
emm_tj_ageclass3 <- emmeans(tj_sla_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_tj_ageclass3)
```

```{r}
if(!requireNamespace("emmeans", quietly = TRUE)) install.packages("emmeans")
if(!requireNamespace("pbkrtest", quietly = TRUE)) install.packages("pbkrtest")


# Calculate estimated marginal means for the interaction
emm_sla_tj <- emmeans(tj_sla_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_tj <- summary(emm_sla_tj) %>%
  as.data.frame()

# Define the position dodge width
dodge_width <- 0.75

# Plotting with boxplots for the raw data
p <- ggplot(tjøtta_data, aes(x = species, y = log(SLA), fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  # Overlay the EMM predictions as points for clarity on the model's fit
  geom_point(data = emm_tj, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, ageClass)),
             position = position_dodge(width = dodge_width), size = 3) +
  # Facet by ageClass to compare across age classes
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "SLA by Treatment, Species, and Age Class",
       x = "Species",
       y = "SLA",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(values = c("Amb (0)" = "lightblue", "Ext (90)" = "salmon")) +
  scale_color_manual(values = c("Amb (0)" = "blue", "Ext (90)" = "red"))

# Display the plot
print(p)

```




**trait 4: mean thickness*

```{r}
tj_t_s_d_a <- lmer(mean_thickness ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = tjøtta_data)
summary(tj_t_s_d_a)


#validate model
performance::check_model(tj_t_s_d_a, detrend = FALSE)
anova(tj_t_s_d_a)

#1. species effefct
emm_tj_species4 <- emmeans(tj_t_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_tj_species4)

#2. drought effect
# assessing the effect of drought across all species and age classes
emm_tj_drought4 <- emmeans(tj_t_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_tj_drought4)

# 3. age effect
# assessing the effect of age class within each species and drought treatment
emm_tj_ageclass4 <- emmeans(tj_t_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_tj_ageclass4)
```


```{r}
if(!requireNamespace("emmeans", quietly = TRUE)) install.packages("emmeans")
if(!requireNamespace("pbkrtest", quietly = TRUE)) install.packages("pbkrtest")


# Calculate estimated marginal means for the interaction
emm_thickness_tj <- emmeans(tj_t_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_tj <- summary(emm_thickness_tj) %>%
  as.data.frame()

# Define the position dodge width
dodge_width <- 0.75

# Plotting with boxplots for the raw data
p <- ggplot(tjøtta_data, aes(x = species, y = mean_thickness, fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  # Overlay the EMM predictions as points for clarity on the model's fit
  # Ensure that the position dodge here matches the dodge for the boxplot
  geom_point(data = emm_tj, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, ageClass)),
             position = position_dodge(width = dodge_width), size = 3) +
  # Facet by ageClass to compare across age classes
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Thickness by Treatment, Species, and Age Class",
       x = "Species",
       y = "Thickness",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(values = c("Amb (0)" = "lightblue", "Ext (90)" = "salmon")) +
  scale_color_manual(values = c("Amb (0)" = "blue", "Ext (90)" = "red"))

# Display the plot
print(p)

```




 trait 5. leaf area
```{r}
tj_area_s_d_a <- lmer(log(leaf_area) ~ DroughtTrt * species * ageClass +
                                    (1 | DroughNet_plotID), 
                                  data = tjøtta_data)
summary(tj_area_s_d_a)


#validate model
performance::check_model(tj_area_s_d_a, detrend = FALSE)
anova(tj_area_s_d_a)

#1. species effefct
emm_tj_species5 <- emmeans(tj_area_s_d_a, pairwise ~ species | DroughtTrt * ageClass, adjust = "tukey")
summary(emm_tj_species5)

#2. drought effect
# assessing the effect of drought across all species and age classes
emm_tj_drought5 <- emmeans(tj_area_s_d_a, pairwise ~ DroughtTrt | species * ageClass, adjust = "tukey")
summary(emm_tj_drought5)

# 3. age effect
# assessing the effect of age class within each species and drought treatment
emm_tj_ageclass5 <- emmeans(tj_area_s_d_a, pairwise ~ ageClass | species * DroughtTrt, adjust = "tukey")
summary(emm_tj_ageclass5)
```


```{r}
if(!requireNamespace("emmeans", quietly = TRUE)) install.packages("emmeans")
if(!requireNamespace("pbkrtest", quietly = TRUE)) install.packages("pbkrtest")


# Calculate estimated marginal means for the interaction
emm_height_tj <- emmeans(tj_area_s_d_a, ~ DroughtTrt * species * ageClass)

# Convert EMMs to a dataframe for plotting
emm_tj <- summary(emm_height_tj) %>%
  as.data.frame()

# Define the position dodge width
dodge_width <- 0.75

# Plotting with boxplots for the raw data
p <- ggplot(tjøtta_data, aes(x = species, y = log(leaf_area), fill = DroughtTrt)) +
  geom_boxplot(position = position_dodge(width = dodge_width)) +
  # Overlay the EMM predictions as points for clarity on the model's fit
  geom_point(data = emm_tj, aes(x = species, y = emmean, color = DroughtTrt, group = interaction(DroughtTrt, ageClass)),
             position = position_dodge(width = dodge_width), size = 3) +
  # Facet by ageClass to compare across age classes
  facet_wrap(~ ageClass, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Leaf Area by Treatment, Species, and Age Class",
       x = "Species",
       y = "Log(Leaf Area)",
       fill = "Drought Treatment", color = "Drought Treatment") +
  scale_fill_manual(values = c("Amb (0)" = "lightblue", "Ext (90)" = "salmon")) +
  scale_color_manual(values = c("Amb (0)" = "blue", "Ext (90)" = "red"))

# Display the plot
print(p)


```


stop here



























**plot model onto the ggplot** keep

```{r}
ggplot(lygra_data, aes(x = species, y = LDMC, fill = DroughtTrt)) +
  geom_boxplot() +
  facet_grid(~ ageClass, scales = "free") +  # Facet only by ageClass
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), # vertical adjustment for x axis labels
        plot.caption = element_text(hjust = 0.5)) + # Center the caption
  labs(title = "LDMC vs Treatment by Species",
       x = "Species",
       y = "LDMC",
       fill = "Drought Treatment",
       caption = "Figure 3: LDMC Variation in Dwarf Shrubs Under Drought Conditions Across Different Successional Phases") +
  scale_fill_manual(values = c("Amb (0)" = "darkblue", "Ext (90)" = "lightblue"))

#fit model

# Get fitted values directly from the model
fitted_values <- predict(LY_LDMC_species_droughtTrt, re.form = NA)  

lygra_data$fitted <- fitted_values

# Assuming fitted_values has been added to filtered_data as shown above
ggplot(lygra_data, aes(x = species, y = LDMC, fill = DroughtTrt)) +
  geom_boxplot() +
  geom_point(aes(y = fitted), position = position_dodge(width = 0.75), color = "red", size = 1.5) + # Add fitted values
  facet_grid(~ ageClass, scales = "free") + # Facet only by ageClass
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), # Vertical adjustment for x axis labels
        plot.caption = element_text(hjust = 0.5)) + 
  labs(title = "LDMC vs Treatment by Species",
       x = "Species",
       y = "LDMC",
       fill = "Drought Treatment",
       caption = "Figure 3: LDMC Variation in Dwarf Shrubs Under Drought Conditions Across Different Successional Phases") +
  scale_fill_manual(values = c("Amb (0)" = "darkblue", "Ext (90)" = "lightblue"))




#fiiting in the geom jitter plot

# Get fitted values directly from the model
fitted_values <- predict(LY_LDMC_species_droughtTrt, re.form = NA)  

# Assuming 'fitted_values' is a vector of the same length as 'lygra_data'
lygra_data$fitted <- fitted_values

fitted_values$dodge_position <- as.numeric(lygra_data$DroughtTrt) - 1.5

ggplot(lygra_data, aes(x = species, y = LDMC, group = DroughtTrt)) +
  geom_jitter(aes(color = DroughtTrt), position = position_dodge(width = 0.5), size = 1.5, alpha = 0.7) +
  geom_point(aes(y = fitted, color = DroughtTrt),  # Ensure 'fitted' is now recognized
             position = position_dodge(width = 0.5), size = 3, shape = 16, color = "black") +
  facet_grid(ageClass ~ siteID, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "LDMC vs Treatment by Species", x = "Species", y = "LDMC", color = "Drought Treatment") +
  scale_color_manual(values = c("Amb (0)" = "darkgreen", "Ext (90)" = "deeppink"))

```









**young and old leaves (lygra)**

```{r}
# Subset the data for the specified species with young and old leaves, and for the Lygra site
subset_data1 <- subset(droughtnet_data2_clean, species %in% c("Vaccinium vitis-idaea", "Empetrum nigrum") & siteID == "Lygra")

# View the first few rows of the subsetted data
View(subset_data1)

```

**LDMC**

```{r}
ldmc_three_way_interactions <- lmer(LDMC ~ species * DroughtTrt * leaf_age +
                                (1 | DroughNet_plotID/plant_nr), 
                              data = subset_data1)

summary(ldmc_three_way_interactions)
performance::check_model(ldmc_three_way_interactions, detrend = FALSE)
```

**SLA**

```{r}
sla_three_way_interactions <- lmer(log(SLA) ~ species * DroughtTrt * leaf_age +
                                (1 | DroughNet_plotID/plant_nr), 
                              data = subset_data1)

summary(sla_three_way_interactions)
performance::check_model(sla_three_way_interactions, detrend = FALSE)
anova(sla_three_way_interactions)
```

**mean thickness**

```{r}
thickness_three_way_interactions <- lmer(mean_thickness ~ species * DroughtTrt * leaf_age +
                                (1 | DroughNet_plotID/plant_nr), 
                              data = subset_data1)

summary(thickness_three_way_interactions)
performance::check_model(thickness_three_way_interactions, detrend = FALSE)
```

**young and old leaves (tjøtta)**

```{r}
# Subset the data for the specified species with young and old leaves, and for the Lygra site
subset_data2 <- subset(droughtnet_data2_clean, species %in% c("Vaccinium vitis-idaea", "Empetrum nigrum") & siteID == "Tjøtta")

# View the first few rows of the subsetted data
View(subset_data2)
```

**LDMC**

```{r}
ldmc1_three_way_interactions <- lmer(LDMC ~ species * DroughtTrt * leaf_age +
                                (1 | DroughNet_plotID/plant_nr), 
                              data = subset_data2)

summary(ldmc1_three_way_interactions)
performance::check_model(ldmc1_three_way_interactions, detrend = FALSE)
```

**SLA**

```{r}
sla2_three_way_interactions <- lmer(log(SLA) ~ species * DroughtTrt * leaf_age +
                                (1 | DroughNet_plotID/plant_nr), 
                              data = subset_data2)

summary(sla2_three_way_interactions)
performance::check_model(sla2_three_way_interactions, detrend = FALSE)
```

**mean thickness**

```{r}
thickness3_three_way_interactions <- lmer(mean_thickness ~ species * DroughtTrt * leaf_age +
                                (1 | DroughNet_plotID/plant_nr), 
                              data = subset_data2)

summary(thickness3_three_way_interactions)
performance::check_model(thickness3_three_way_interactions, detrend = FALSE)
```

**CALLUNA SHOOT TYPE (lygra)**

```{r}
# Create a data frame for Calluna vulgaris at the Lygra site and remove rows with NA values
calluna_shoot <- droughtnet_data2_clean %>%
  filter(species == "Calluna vulgaris" & siteID == "Lygra") %>%
  drop_na()

view(calluna_shoot)
```

**LDMC**

```{r}
shoot_type_lmdc3 <- lmer(LDMC ~ (calluna_shoot_type * DroughtTrt * ageClass) +
                               (1 | DroughNet_plotID/plant_nr), 
                             data = calluna_shoot)
summary(shoot_type_lmdc3)

performance::check_model(shoot_type_lmdc3, detrend = FALSE)
```

**SLA**

```{r}
shoot_type_sla3 <- lmer(log(SLA) ~ (calluna_shoot_type * DroughtTrt * ageClass) +
                               (1 | DroughNet_plotID/plant_nr), 
                             data = calluna_shoot)
summary(shoot_type_sla3)

performance::check_model(shoot_type_sla3, detrend = FALSE)
```

**mean thickness**

```{r}
shoot_type_thickness3 <- lmer(mean_thickness ~ (calluna_shoot_type * DroughtTrt * ageClass) +
                               (1 | DroughNet_plotID/plant_nr), 
                             data = calluna_shoot)
summary(shoot_type_thickness3)

performance::check_model(shoot_type_thickness3, detrend = FALSE)
```
