---
title: "Calluna data cleaning"
output: html_notebook
---


Load the Data.,

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(readr)

data <- read_csv("DURIN_clean.csv")
View(data)

```
Filter data for species "Calluna vulgaris"
```{r}
calluna_data <- filter(data, species == "Calluna vulgaris" & siteID %in% c("Lygra", "Sogndal", "Senja"))
view(calluna_data)
 ```
# Count the number of "Calluna vulgaris" plants per site
(plant_count_per_site <- calluna_data %>%
  filter(siteID %in% c("Lygra", "Sogndal", "Senja")) %>%
  group_by(siteID) %>%
  summarise(count = n()))
```

species count (number of Calluna plants) per plot
```{r}
count_per_plot <- calluna_data %>%
  group_by(DURIN_plot) %>%
  summarise(count = n())
view(count_per_plot)

count_per_plot <- calluna_data %>%
  group_by(DroughNet_plotID) %>%
  summarise(count = n())
view(count_per_plot)
```

no plots assigned
```{r}
# Count plants with no plots assigned
no_plot_count <- calluna_data %>%
  filter(is.na(DURIN_plot) & is.na(DroughNet_plotID)) %>%
  summarise(count = n()) %>%
  mutate(plot_type = "No Plot Assigned", plot = "No Plot Assigned")

# Combine the results
combined_results <- bind_rows(count_durin_plot, count_droughnet_plot, no_plot_count)

 view(combined_results)

# Filter out all entries where both DURIN_plot and DroughNet_plotID are not assigned
no_plots_assigned_by_envelope <- calluna_data %>%
  filter(is.na(DURIN_plot) & is.na(DroughNet_plotID)) %>%
  select(envelope_ID, siteID, DURIN_plot, DroughNet_plotID)

view(no_plots_assigned_by_envelope)
```
#24 plants do not have plot assigned
#solution- did a crosscheck with the main data set, all of them belong to phys

check multiple heights for calluna vulgaris
```{r}
  # Filter for Calluna vulgaris, exclude NA in DURIN_plot, group by plant_nr and plant_height, and gather envelope_IDs
calluna_heights_and_envelopes <- calluna_data %>%
  filter(!is.na(DURIN_plot)) %>%  # Exclude rows where DURIN_plot is NA
  group_by(siteID, DURIN_plot, plant_nr, plant_height) %>%
  summarise(
    envelope_count = n(),
    envelope_ids = list(envelope_ID)
  ) %>%
  ungroup()  # Make sure to ungroup after summarise

# Filter to only include rows where there's more than one height for the same plant
multiple_heights_calluna <- calluna_heights_and_envelopes %>%
  group_by(siteID, DURIN_plot, plant_nr) %>%
  filter(n() > 1) %>%
  ungroup()

# If you want to view this in R console
view(multiple_heights_calluna)

# Expand the envelope_ID column to display all the barcodes
expanded_calluna <- multiple_heights_calluna %>%
  unnest(envelope_ids)

# If you want to view this in R console
view(expanded_calluna)
#found a good number of them with multple heights
#solution- to crosscheck
```

plants with missing heights
```{r}
# Filter data for the three sites and missing heights
calluna_missing_heights <- calluna_data %>%
  group_by(envelope_ID) %>%
  filter(siteID %in% c("Lygra", "Sogndal", "Senja"), is.na(plant_height))

# Display the plants with missing heights
view(calluna_missing_heights %>% select(envelope_ID, siteID, DURIN_plot, DroughNet_plotID, plant_height))
```
#29 calluna missing height
#solution-did a cross check with main data set, all belong to phys

missing shoot type
```{r}
missing_shoot_type<- calluna_data %>%
  group_by(envelope_ID) %>%
  filter(!calluna_shoot_type %in% c("Short", "Long"))

view(missing_shoot_type %>% select(envelope_ID, calluna_shoot_type))

unique(calluna_data$calluna_shoot_type)
#Could you please review this code? Instead of filtering out entries without a shoot type, the resulting table seems to have a mix of both entries with and without shoot types, dont know where am missing out in this

```


```
